Making Sequential Programs Deterministic

### 2019-04-06 introduction.tex
intro citations

A nondeterministic program may produce different output on different runs
when provided with the same input.
Nondeterminism is a serious problem for software developers and users.
\todo{Add lots of citations in the below.}
\begin{itemize}
\item
  Nondeterminism makes a program difficult to \textbf{test}, because test
  oracles must account for all possible behaviors while still enforcing
  correct behaviors.  Test oracles that are too strict lead to flaky
% ,Gyori:2015:RTD:2771783.2771793
  tests~\cite{LuoHEM2014,nondex,deflaker,Rahman:2018:IFF:3236024.3275529}
  that sometimes pass and sometimes fail.  Flaky tests must be re-run, or
  developers ignore them; in either case, their utility to detect defects
  is limited.
\item
  Nondeterminism makes it difficult to \textbf{compare} two runs of a
  program on different data, or to compare a run of a slightly modified
  program to an original program.  This hinders debugging and maintenance,
  and prevents use of techniques such as Delta Debugging~\cite{Zeller1999}.
%\item
%  Nondeterminism may make outputs incompatible with previous outputs relied
%  on by users or by other systems~\todo{These do not seem to be supporting
%    the claim about bugs related to incompatibilities between outputs.
%    These seem to be general citations about test flakiness or dependence.
%    Readers will be unhappy if we make a claim but then do not support it.}\cite{Herzig:2015:EDF:2819009.2819018,Huo:2014:IOQ:2635868.2635917,ZhangJWMLEN2014,BellKMD2015,Fowler,Gyori:2015:RTD:2771783.2771793,Dan:2013:10.1007/978-3-642-39038-8_25,Vahabzadeh:2015:7332456,Sudarshan}.
%  \todo{Citations about bugs related to this?}
\item
  Nondeterminism reduces users' and developers' \textbf{trust} in a program's output.
\end{itemize}

@@

The high-level goal of our work is to provide programmers with a tool for
specifying deterministic properties in a program and verifying them
statically.
%
Other researchers have also recognized the importance of the problem of nondeterminism.
Previous work in program analysis for nondeterminism has focused on unsound dynamic
approaches that identify flaky test cases.
NonDex~\cite{nondex} uses a modified JVM that returns different results on different
executions, for a few key JDK methods with loose specifications.  Running a
test suite multiple times can reveal unwarranted dependence on those
methods.
%NonDex~\cite{nondex} is a tool that manually\todo{This is confusing:  the
%  tool can't manually identify.  Do you mean that the developers manually
%  identified, then the tool uses those models?}
%identifies 
DeFlaker~\cite{deflaker} looks at a range of commit versions
of a code, and marks a test as flaky if it doesn't execute any modified code but still fails in the newer version. These techniques
have been able to identify issues in real-world programs, some of which
have been fixed by the developers. We believe that identifying and
resolving nondeterminism
at an earlier phase in the software development lifecycle would be greatly beneficial to
developers.

>>>>

A nondeterministic program may produce different output on different runs
when provided with the same input.
Nondeterminism is a serious problem for software developers and users.
\todo{Add lots of citations in the below.}
\begin{itemize}
\item
  Nondeterminism makes a program difficult to \textbf{test}, because test
  oracles must account for all possible behaviors while still enforcing
  correct behaviors.  Test oracles that are too strict lead to flaky
% ,Gyori:2015:RTD:2771783.2771793
  tests~\cite{LuoHEM2014,nondex,deflaker,Rahman:2018:IFF:3236024.3275529,LuoHEM2014,Sudarshan}
  that sometimes pass and sometimes fail.  Flaky tests must be re-run, or
  developers ignore them; in either case, their utility to detect defects
  is limited.
\item
  Nondeterminism makes it difficult to \textbf{compare} two runs of a
  program on different data, or to compare a run of a slightly modified
  program to an original program.  This hinders debugging and maintenance,
  and prevents use of techniques such as Delta Debugging~\cite{Zeller1999,YuLCZ2012}.
%\item
%  Nondeterminism may make outputs incompatible with previous outputs relied
%  on by users or by other systems~\todo{These do not seem to be supporting
%    the claim about bugs related to incompatibilities between outputs.
%    These seem to be general citations about test flakiness or dependence.
%    Readers will be unhappy if we make a claim but then do not support it.}\cite{Herzig:2015:EDF:2819009.2819018,Huo:2014:IOQ:2635868.2635917,ZhangJWMLEN2014,BellKMD2015,Fowler,Gyori:2015:RTD:2771783.2771793,Dan:2013:10.1007/978-3-642-39038-8_25,Vahabzadeh:2015:7332456,Sudarshan}.
%  \todo{Citations about bugs related to this?}
\item
  Nondeterminism reduces users' and developers' \textbf{trust} in a program's output.
\end{itemize}

@@

The high-level goal of our work is to provide programmers with a tool for
specifying deterministic properties in a program and verifying them
statically.
%
Other researchers have also recognized the importance of the problem of nondeterminism.
Previous work in program analysis for nondeterminism has focused on unsound dynamic
approaches that identify flaky test cases.
NonDex~\cite{nondex} uses a modified JVM that returns different results on different
executions, for a few key JDK methods with loose specifications.  Running a
test suite multiple times can reveal unwarranted dependence on those
methods.
%NonDex~\cite{nondex} is a tool that manually\todo{This is confusing:  the
%  tool can't manually identify.  Do you mean that the developers manually
%  identified, then the tool uses those models?}
%identifies 
DeFlaker~\cite{deflaker} looks at a range of commit versions
of a code, and marks a test as flaky if it doesn't execute any modified code but still fails in the newer version. These techniques
have been able to identify issues in real-world programs, some of which
have been fixed by the developers. We believe that identifying and
resolving nondeterminism
at an earlier phase in the software development lifecycle would be greatly beneficial to
developers~\cite{briski2008minimizing}.
